Dynamical Properties of Molecular Cloud Complexes in a Redshift 6 Prototypical Galaxy at the Epoch of Reionization

### 2018-09-11 WRITEUP/ms.tex
more polishing and add ref

For visualization, we overplot the molecular structures identified using
a subset of the $n_{\rm cut}$ (0.53, 6.81, and 19.00\,\cc) on the H$_2$ density maps 
in \Fig{MC} as an example.
The plots shown in \Fig{MC} are weighted by the total gas density; that is, 
\begin{equation}
{\rm pixel} = \frac{\sum_i~\rho_{\rm H2, i}~w_i~\Delta z_i}{\sum_i~w_i~\Delta z_i}, 
\label{eqn:integrate}
\end{equation} 
where $\rho_{\rm H2, i}$ is the H$_2$ density of pixel $i$, $w_i$ is the weighted quantity in pixel $i$ which we adopt 
to be the total gas density, and $\Delta z_i$ is the path length of pixel $i$ along the line-of-sight over which the quantities are integrated along.

@@

In the following subsections,
we examine the MC properties in the context of the Larson's relations \citep{Larson81a}, which
describe interactions between gravity and turbulence and is the first set of 
relations used for studying star-forming molecular structure formation based on observables (linewidth-size,
density-size, and mass-size relations) and for comparing properties of molecular structures
in different galactic environments.
% Inner MW: R=2-8.5 kpc, most of the mass in the molecular ISM is in the form of GMCs with sizes $gtrsim$20\,pc and masses greater than 10$^5$\,\Msun \citep{Sanders85a}.
Observations of nearby gas-rich galaxies such as M64 and NGC\,253
show higher velocity dispersions compared
to the disk/mid-plane of the Milky Way which are, on the other hand,
more consistent with those observed in the inner regions of
Milky Way \citep{Oka01a, Rosolowsky05a, Heyer09a, Leroy15a}.  % also higher $P_{\rm int}$,
% which is $\propto\rho\sigma^2$
While the velocity dispersions of the clouds in these nearby gas-rich galaxies are comparable
to the inner regions of Milky Way, they are found to be bigger in size (approximately
an order of magnitude) and mass (approximately two orders of magnitude).
These bigger clouds have been interpreted as a results of their higher gas mass fractions and
surface densities and velocity dispersions, since
fragmentation occurs near the Jeans length $L_J\propto\sigma^2/\Sigma$
and Jeans mass $M_J\propto\sigma^4/\Sigma$
for dispersion-supported structures.

@@

While it may appear that there is an inconsistency between the high $\alpha_{\rm vir}$
nature of the majority of the molecular gas in \flower and its SFR.
That is, if most of the molecular gas in \flower has high $\alpha_{\rm vir}$, why does it sustain its high SFR of $\sim$100\,\Msun\,yr\pmOne?
This discrepancy can be explained in two ways.
First, our results are limited by the resolution of the simulation, such that, in each MC, there
likely exist multiple smaller-scales molecular structures (e.g., clumps and cores),
as in the classical MC hierarchy. These smaller structures are 
no longer supported by large-scale gravitational potential and differential rotation,
as turbulence has dissipated rapidly on these smaller scales to enable collapse \citep{Clark04a}.
In addition, non-axisymmetric perturbations (e.g., arms) to the gravitational potential
can induce orbit crossing, shocks and dissipation in gas, promoting turbulence dissipation.

@@

% interpretation
Velocity dispersion and surface density are tightly connected to pressure. 
That is, the pressure of an MC is expected to increase with its velocity dispersion and surface density (see e.g., \Eq{pext} and 
since $P\propto\Sigma^2$).
In the central molecular zone of the Milky Way (CMZ), we have seen molecular clouds with high pressures of 
$P/k_B > $10$^7$\,K\,cm$^{-3}$. Such high pressures are also seen in massive star-forming regions such as 30 Doradus.
The higher $\Sigma_{\rm gas}$ in turn is related to more massive clouds being formed.
In M64, there are a few clouds with masses $M_{\rm cl}>$10$^7$\,\Msun. 
On the other hand, such massive molecular structures are unknown to local group, and are likely the sites 
from which massive star clusters or OB associations are to be formed.
% 
The shallower slope found in the CMF for \flower compared to those observed in the nearby Universe (see \Sec{cmf}) may
suggest that the MCs of \flower are more massive compared to those observed in the nearby Universe,
notwithstanding the limited numerical resolution of our simulation.
Together with the high $P_{\rm ext}$ may suggest different IMF for \flower compared to those observed in nearby galaxies.
This could be understood in the context of .... different cooling properties will lead to different xxx, and thus, Jeans mass/thus determines the final properties of the cloud fragmentation....  
% (While the first stars and galaxies are cooled predominated via atomic H and H$_2$ lines.
%As shown in the $T-n$ phase plot in Fig. 8 of \citealt{Pallottini17b}, 
%metals also play an important role in the cooling of the first galaxies soon after the first generation of stars have enriched the surrounding.)

@@

We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy 
at the EoR using state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by metals, and stellar feedback.
We use a clump finder algorithm and a set of H$_2$ volumetric densities
to identify MCs and their sub-structures in the main galaxy --- \flower\ --- of the
simulated 20\,Mpc h\pmOne box and in its satellites.
We decompose the molecular structures into non-overlapping tiles
by identifying a set of different density contours at different snapshots.
Using volumetric H$_2$ is essentially similar to identifying molecular structures using
different contour levels on a surface brightness map of a molecular line tracer (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).
% Our choice of density cut is analogues to the .... in \obs. It should be ... that both suffer from problem .... For instance, in \obs.. limited S/N.
We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter for each MC and 
compare them with those observed in the Milky Way disk, the Galactic center,
and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF.
We also examine their properties at the different evolutionary stages of \flower and compare
them with observations.

@@

We compare the dynamics of MCs in \flower to \obs in the context of the Larson's relation.
The MCs of \flower are found to have higher $\sigma$ and $\Sigma$ systematically regardless of the $n_{\rm cut}$ adopted.
Our results are thus insensitive to the various density cuts of choice.
%However, our results are dependent on the numerical resolution of the simulation.
The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the $n_{\rm cut}$ and even for  
the molecular substructures.
Such a high velocity dispersion likely results from 
the strong supernova and stellar feedback \flower experienced over the multiple episodes of bursty \SF.
Virial analysis indicates that the MC/arms of the main disk of \flower are unbound, but the substructures have lower virial parameter. 
This is consistent with the notion that collapsing structures result from gravitational instability within globally stable structures, which are 
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as 
result of the weaker stellar feedback as they have
experienced less episodes of \SF compared to \flower (also supported by the lower stellar-to-gas mass ratio of the latter).
MCs in the satellites are therefore likely collapsing structures. This paints a picture, in which at the EoR, \SF continues 
as gas is being accreted from the IGM.
We find no temporal variations in the MC dynamics over the course of 700\,Myr traced in our simulation, at least in terms of the scaling relations examined. 
Our results are independent of the volume density threshold adopted, except for the slope of the cumulative 
mass distribution, which steepens as we increase $n_{\rm cut}$.

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatio-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.

High resolution zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and is subject to the input model/sub-grid models adopted, 
provide an avenue to examine and postulate/make testable predictions on the morphology and dynamics of
the multi-phase ISM structures of the first galaxies and their satellite galaxies.
High resolution imaging of the molecular gas content in the first galaxies with ALMA and the ngVLA, on the other hand,
have the potential to provide useful observational data to test our findings and the validity of our simulation to shed
light on \SF since the cosmic dark ages.

% model can guide and interpret observations and provide new testable predictions. 
% refine models using these new observations and test where models fail.
% The opposite is true, where can use models to make predictions for the first galaxies, 
% where future obs w/ ALMA and ngVLA will help discriminate and better understand xxx.

@@

We thank Jens Kauffmann, Thushara Pillai, and Mark Swinbank for sharing their data.
T.K.D.L. acknowledges support by the NSF through award SOSPA4-009
from the NRAO and support from the Simons Foundation.
A.F. acknowledges support from the ERC Advanced Grant INTERSTELLAR H2020/740120.
This work was initiated as a project for the Kavli Summer Program in Astrophysics (KSPA)
held at the Center for
Computational Astrophysics of the Flatiron Institute in 2018. The program was co-funded by the Kavli
Foundation and the Simons Foundation.
We thank the KSPA Scientific and Local Organizing Committees, and the program founder,
Pascale Garaud for supporting the genesis of this work.
We also thank the New York University CCPP for their hospitality in hosting us, the refugees, after the steam pipe explosion in NYC during the KSPA.

>>>>

For visualization, we overplot the molecular structures identified using
a subset of the $n_{\rm cut}$ (0.53, 6.81, and 19.00\,\cc) on the H$_2$ density maps 
in \Fig{MC} as an example.
The plots shown in \Fig{MC} are weighted by the total gas density; that is, 
\begin{equation}
{\rm voxel_i} = \frac{\sum_i~\rho_{\rm H2, i}~w_i~\Delta z_i}{\sum_i~w_i~\Delta z_i}, 
\label{eqn:integrate}
\end{equation} 
where $\rho_{\rm H2, i}$ is the H$_2$ density of voxel $i$, $w_i$ is the weighted quantity in voxel $i$ which we adopt 
to be the total gas density, and $\Delta z_i$ is the path length of voxel $i$ along the line-of-sight over which the quantities are integrated along.

@@

In the following subsections,
we examine the MC properties in the context of the Larson's relations \citep{Larson81a}, which
describe interactions between gravity and turbulence and is the first set of 
relations used for studying star-forming molecular structure formation based on observables (linewidth-size,
density-size, and mass-size relations) and for comparing properties of molecular structures
in different galactic environments.
% Inner MW: R=2-8.5 kpc, most of the mass in the molecular ISM is in the form of GMCs with sizes $gtrsim$20\,pc and masses greater than 10$^5$\,\Msun \citep{Sanders85a}.
Observations of nearby gas-rich galaxies such as M64 and NGC\,253
show higher velocity dispersions compared
to the disk/mid-plane of the Milky Way which are, on the other hand,
more consistent with those observed in the inner regions of
Milky Way and M51 \citep{Oka01a, Rosolowsky05a, Heyer09a, Hughes13b, Leroy15a, Rice16a}.  % also higher $P_{\rm int}$,
% which is $\propto\rho\sigma^2$
% M51: Colombo+14
While the velocity dispersions of the clouds in these nearby gas-rich galaxies are comparable
to the inner regions of Milky Way, they are found to be bigger in size (approximately
an order of magnitude) and mass (approximately two orders of magnitude).
These bigger clouds have been interpreted as a results of their higher gas mass fractions and
surface densities and velocity dispersions, since
fragmentation occurs near the Jeans length $L_J\propto\sigma^2/\Sigma$
and Jeans mass $M_J\propto\sigma^4/\Sigma$
for dispersion-supported structures.

@@

While it may appear that there is an inconsistency between the high $\alpha_{\rm vir}$
nature of the majority of the molecular gas in \flower and its SFR.
That is, if most of the molecular gas in \flower has high $\alpha_{\rm vir}$, why does it sustain its high SFR of $\sim$100\,\Msun\,yr\pmOne?
This discrepancy can be explained in two ways.
First, our results are limited by the resolution of the simulation, such that, in each MC, there
likely exist multiple smaller-scales molecular structures (e.g., clumps and cores),
as in the classical MC hierarchy. 
Second, due to 
turbulence dissipation on small-scales, such (sub-)structures are no longer supported by 
large-scale gravitational potential and differential rotation. 
This in turn enables subregions of the MCs to collapse and potentially form OB associations out of these unbound
MCs \citep{Clark04a, Clark05a}.
Additional mechanisms causing non-axisymmetric perturbations (e.g., arms) to the gravitational potential
can also induce orbit crossing, shocks and dissipation in gas, further promoting turbulence dissipation.

@@

% interpretation
Velocity dispersion and surface density are tightly connected to pressure. 
That is, the pressure of an MC is expected to increase with its velocity dispersion and surface density (see e.g., \Eq{pext} and 
since $P\propto\Sigma^2$).
In the central molecular zone of the Milky Way (CMZ), we have seen molecular clouds with high pressures of 
$P/k_B > $10$^7$\,K\,cm$^{-3}$. Such high pressures are also seen in massive star-forming regions such as 30 Doradus.
The higher $\Sigma_{\rm gas}$ in turn is related to more massive clouds being formed.
In M64 and M51, there are a few clouds with masses $M_{\rm cl}>$10$^7$\,\Msun \citep{Rosolowsky05a, Koda09a, Colombo14a}. 
On the other hand, such massive molecular structures are unknown to Local Group \citep[e.g.,][]{Rice16a}, and are likely the sites 
from which massive star clusters or OB associations are to be formed.
% 
The shallower slope found in the CMF for \flower compared to those observed in the nearby Universe (see \Sec{cmf}) may
suggest that the MCs of \flower are more massive compared to those observed in the nearby Universe,
notwithstanding the limited numerical resolution of our simulation.
Together with the high $P_{\rm ext}$ may suggest different IMF for \flower compared to those observed in nearby galaxies.
This could be understood in the context of .... different cooling properties will lead to different xxx, and thus, Jeans mass/thus determines the final properties of the cloud fragmentation....  
% (While the first stars and galaxies are cooled predominated via atomic H and H$_2$ lines.
%As shown in the $T-n$ phase plot in Fig. 8 of \citealt{Pallottini17b}, 
%metals also play an important role in the cooling of the first galaxies soon after the first generation of stars have enriched the surrounding.)

@@

We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy 
at the EoR using state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by metals, and stellar feedback.
We use a clump finder algorithm and a set of H$_2$ volumetric densities
to identify MCs and their sub-structures in the main galaxy --- \flower\ --- of the
simulated 20\,Mpc h\pmOne box and in its satellites.
We decompose the molecular structures into non-overlapping tiles
by identifying a set of different density contours at different snapshots.
Using volumetric H$_2$ is essentially similar to identifying molecular structures using
different contour levels/surfaces on a surface brightness map/cube of a molecular line tracer (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).
% Our choice of density cut is analogues to the .... in \obs. It should be ... that both suffer from problem .... For instance, in \obs.. limited S/N.
We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter for each MC and 
compare them with those observed in the Milky Way disk, the Galactic center,
and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF.
We also examine their properties at the different evolutionary stages of \flower and compare
them with observations.

@@

We compare the dynamics of MCs in \flower to \obs in the context of the Larson's relation.
The MCs of \flower are found to have higher $\sigma$ and $\Sigma$ systematically regardless of the $n_{\rm cut}$ adopted.
Our results are thus insensitive to the various density cuts of choice.
%However, our results are dependent on the numerical resolution of the simulation.
The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the $n_{\rm cut}$ and even for  
the molecular substructures.
Such a high velocity dispersion likely results from 
the strong supernova and stellar feedback \flower experienced over the multiple episodes of bursty \SF.
Virial analysis indicates that the MC/arms of the main disk of \flower are unbound, but the substructures have lower virial parameter. 
This is consistent with the notion that collapsing structures result from gravitational instability within globally stable structures, which are 
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as 
result of the weaker stellar feedback as they have
experienced less episodes of \SF compared to \flower (also supported by the lower stellar-to-gas mass ratio of the latter).
MCs in the satellites are therefore likely collapsing structures. This paints a picture, in which at the EoR, \SF continues 
as gas is being accreted from the IGM.

We find no temporal variations in the MC dynamics over the course of 700\,Myr traced in our simulation, at least in terms of the scaling relations examined. 
Our results are independent of the volume density threshold adopted, except for the slope of the cumulative 
mass distribution, which steepens as we increase $n_{\rm cut}$.

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
High resolution zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and is subject to the input model/sub-grid models adopted, 
provide an avenue to examine and postulate/make testable predictions on the morphology and dynamics of
the multi-phase ISM structures of the first galaxies and their satellite galaxies.
High resolution imaging of the molecular gas content in the first galaxies with ALMA and the ngVLA, on the other hand,
have the potential to provide useful observational data to test our findings and the validity of our simulation to shed
light on \SF since the cosmic dark ages.

@@

We thank Jens Kauffmann, Thushara Pillai, and Mark Swinbank for sharing their data.
T.K.D.L. acknowledges support by the NSF through award SOSPA4-009
from the NRAO and support from the Simons Foundation.
A.F. acknowledges support from the ERC Advanced Grant INTERSTELLAR H2020/740120.
This work is based on a project developed at the Kavli Summer Program in Astrophysics (KSPA)
held at the Center for
Computational Astrophysics of the Flatiron Institute in 2018. The program was co-funded by the Kavli
Foundation and the Simons Foundation.
We thank the KSPA Scientific and Local Organizing Committees, and the program founder,
Pascale Garaud, for supporting the genesis of this work.
We also thank the Center for Cosmology and Particle Physics (CCPP) 
at the New York University
for their hospitality in hosting us after the steam pipe explosion in NYC during the KSPA.
This research has made use of NASA's Astrophysics Data System Bibliographic Services
and \ncode{yt}, a community-developed Python package for analyzing simulation data in Astronomy.
